---
title: "Assignment 2: Spatial Analysis and Visualization"
subtitle: "Healthcare Access and Equity in Pennsylvania"
author: "Tessa Vu"
date: 2025-10-14
format: 
  html:
    code-fold: false
    toc: true
    toc-location: left
    theme: lux
    embed-resources: true
execute:
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

## Assignment Overview

**Learning Objectives:**

-   Apply spatial operations to answer policy-relevant research
    questions

-   Integrate census demographic data with spatial analysis

-   Create publication-quality visualizations and maps

-   Work with spatial data from multiple sources

-   Communicate findings effectively for policy audiences

------------------------------------------------------------------------

## Part 1: Healthcare Access for Vulnerable Populations

### Research Question

**Which Pennsylvania counties have the highest proportion of vulnerable
populations (elderly + low-income) living far from hospitals?**

Your analysis should identify counties that should be priorities for
healthcare investment and policy intervention.

### Required Analysis Steps

Complete the following analysis, documenting each step with code and
brief explanations:

#### Step 1: Data Collection (5 points)

Load the required spatial data:

-   Pennsylvania county boundaries

-   Pennsylvania hospitals (from lecture data)

-   Pennsylvania census tracts

**Your Task:**

```{r}
# Load required packages
library(kableExtra)
library(knitr)
library(patchwork)
library(scales)
library(sf)
library(tibble)
library(tidycensus)
library(tidyverse)
library(tigris)
library(units)

# Set Census API key
census_api_key("3aaee31789e10b674a531e9f236c35d5394b19ed")

# Load spatial data
pa_counties <- st_read("data/Pennsylvania_County_Boundaries.shp")
pa_tracts <- tracts(state = "PA", cb = TRUE)
pa_hospitals <- st_read("data/hospitals.geojson")

# Check that all data loaded correctly
glimpse(pa_counties)
glimpse(pa_tracts)
glimpse(pa_hospitals)
```

**Questions to answer:**

-   How many hospitals are in your dataset?

    -   223 hospitals.

-   How many census tracts?

    -   3,445 census tracts.

-   What coordinate reference system is each dataset in?

    -   PA county boundaries CRS: GCS WGS 84 (Pseudo-Mercator) / EPSG

        4326. 

    -   PA census tracts CRS: PCS NAD 83 / EPSG 4269.

    -   PA hospitals CRS: GCS WGS 84 (Pseudo-Mercator) / EPSG 4326.

------------------------------------------------------------------------

#### Step 2: Get Demographic Data

Use `tidycensus` to download tract-level demographic data for
Pennsylvania.

**Required variables:**

-   Total population

-   Median household income

-   Population 65 years and over (you may need to sum multiple age
    categories)

**Your Task:**

```{r}
# Get demographic data from ACS
tract_ages <- get_acs(
  geography = "tract",
  variables = c(
    "B01001_020E", "B01001_021E", "B01001_022E",
    "B01001_023E", "B01001_024E", "B01001_025E",
    "B01001_044E", "B01001_045E", "B01001_046E",
    "B01001_047E", "B01001_048E", "B01001_049E"
  ),
  state = "PA",
  year = 2023,
  output = "wide"
)

tract_demographics <- get_acs(
  geography = "tract",
  variables = c(
    total_pop = "B01003_001",
    median_income = "B19013_001"
  ),
  state = "PA",
  year = 2023,
  output = "wide"
)

tract_demographics$age_65_overE <-
  tract_ages$B01001_020E + tract_ages$B01001_021E + tract_ages$B01001_022E +
  tract_ages$B01001_023E + tract_ages$B01001_024E + tract_ages$B01001_025E +
  tract_ages$B01001_044E + tract_ages$B01001_045E + tract_ages$B01001_046E +
  tract_ages$B01001_047E + tract_ages$B01001_048E + tract_ages$B01001_049E
  
tract_demographics$age_65_overM <-
  tract_ages$B01001_020M + tract_ages$B01001_021M + tract_ages$B01001_022M +
  tract_ages$B01001_023M + tract_ages$B01001_024M + tract_ages$B01001_025M +
  tract_ages$B01001_044M + tract_ages$B01001_045M + tract_ages$B01001_046M +
  tract_ages$B01001_047M + tract_ages$B01001_048M + tract_ages$B01001_049M

tract_demographics$TRACT <- NULL

tract_demographics <- rename(tract_demographics, TRACT = NAME)

# Join to tract boundaries
pa_tracts <- left_join(pa_tracts, tract_demographics, by = "GEOID")

summary(pa_tracts)
```

**Questions to answer:**

-   What year of ACS data are you using?

    -   2023 ACS data.

-   How many tracts have missing income data?

    -   65 tracts are missing income estimates and 73 tracts are missing
        income MOEs.

-   What is the median income across all PA census tracts?

    -   \$72,944.

------------------------------------------------------------------------

#### Step 3: Define Vulnerable Populations

Identify census tracts with vulnerable populations based on TWO
criteria:

1.  Low median household income (choose an appropriate threshold)

2.  Significant elderly population (choose an appropriate threshold)

**Your Task:**

```{r}
# Filter for vulnerable tracts based on your criteria
pa_tracts$pct_elderly <- (pa_tracts$age_65_overE / pa_tracts$total_popE) * 100

pa_tracts <- pa_tracts %>%
  mutate(
    low_income = case_when(median_incomeE <= 43700 ~ TRUE,
                           median_incomeE > 43700 ~ FALSE),
    high_elderly = case_when(pct_elderly >= 23.51 ~ TRUE,
                             pct_elderly < 23.51 ~ FALSE),
    vulnerable = case_when(low_income == TRUE | high_elderly == TRUE ~ TRUE,
                           low_income == FALSE & high_elderly == FALSE ~ FALSE)
  )

vulnerable_summary <- pa_tracts %>%
  group_by(vulnerable) %>%
  summarize(
    "Number of Tracts" = n(),
    "Median Income" = median(median_incomeE, na.rm = TRUE),
    "Percent Elderly" = median(pct_elderly, na.rm = TRUE),
    "Percent Vulnerable" = (n() / 3445) * 100
  )

vulnerable_summary
```

**Questions to answer:**

-   What income threshold did you choose and why?

    -   \$43,700 was the income threshold I chose based off of a [2023
        paper published by the U.S. Department of Housing and Urban
        Development’s Office of Policy Development and
        Research](https://www.huduser.gov/portal/datasets/home-datasets/files/HOME_IncomeLmts_State_PA_2023.pdf)
        with a list of PA counties and low income estimates commensurate
        with the area and number of individuals in a household. I took
        the median of all the estimates for the different counties.

-   What elderly population threshold did you choose and why?

    -   I used 23.51% as the threshold because that's the third quartile
        for the elderly population percentage in PA.

-   How many tracts meet your vulnerability criteria?

    -   1,152 tracts meet the vulnerability criteria.

-   What percentage of PA census tracts are considered vulnerable by
    your definition?

    -   Approximately 33.44% of PA's census tracts are considered
        vulnerable.

------------------------------------------------------------------------

#### Step 4: Calculate Distance to Hospitals

For each vulnerable tract, calculate the distance to the nearest
hospital.

**Your Task:**

```{r}
vulnerable_tracts <- pa_tracts %>%
  filter(vulnerable == TRUE)

# Transform to appropriate projected CRS
vulnerable_tracts <- st_transform(vulnerable_tracts, crs = 5070)
pa_tracts <- st_transform(pa_tracts, crs = 5070)
pa_hospitals <- st_transform(pa_hospitals, crs = 5070)

# Calculate distance from each tract centroid to nearest hospital
vulnerable_centroids <- st_centroid(vulnerable_tracts)

nearest_hospital_index <- st_nearest_feature(vulnerable_centroids, pa_hospitals)

nearest_hospital_point <- pa_hospitals[nearest_hospital_index,]

hospital_distance <- st_distance(vulnerable_centroids, nearest_hospital_point, by_element = TRUE)
vulnerable_tracts$nearest_hospital_mi <- set_units(hospital_distance, "mi")

summary(vulnerable_tracts)

sum(vulnerable_tracts$nearest_hospital_mi > set_units(15, "mi"))
```

**Requirements:**

-   Use an appropriate projected coordinate system for Pennsylvania

-   Calculate distances in miles

-   Explain why you chose your projection

    -   I chose Albers Equal Area because it was appropriate when
        calculating hospital distances all over the state, if I'd chosen
        a PCS that was more localized like State Planes, then depending
        whichever cardinal direction I selected, it'd have some
        distortion in other regions of Pennsylvania.

**Questions to answer:**

-   What is the average distance to the nearest hospital for vulnerable
    tracts?

    -   4.35816 miles.

-   What is the maximum distance?

    -   28.98276 miles.

-   How many vulnerable tracts are more than 15 miles from the nearest
    hospital?

    -   34 census tracts of the 1,152 total vulnerable tracts are more
        than 15 miles from the nearest hospital.

------------------------------------------------------------------------

#### Step 5: Identify Underserved Areas

Define "underserved" as vulnerable tracts that are more than 15 miles
from the nearest hospital.

**Your Task:**

```{r}
# Create underserved variable
underserved_tracts <- vulnerable_tracts %>%
  filter(nearest_hospital_mi > set_units(15, "mi"))

vulnerable_tracts <- vulnerable_tracts %>% mutate(
  underserved = case_when(
    vulnerable_tracts$nearest_hospital_mi > set_units(15, "mi") ~ TRUE,
    TRUE ~ FALSE))

(nrow(underserved_tracts) / nrow(vulnerable_tracts)) * 100

glimpse(underserved_tracts)
glimpse(vulnerable_tracts)
```

**Questions to answer:**

-   How many tracts are underserved?

    -   34 census tracts are underserved.

-   What percentage of vulnerable tracts are underserved?

    -   2.951389% of vulnerable census tracts are underserved.

-   Does this surprise you? Why or why not?

    -   Not quite. I'm from Utah and the disparity with hospital access
        is expected, especially since it's twice the size of
        Pennsylvania. I know in the US' East Coast I expect that the
        states have better hospital coverage given that they were also
        established as states over a century earlier than Utah, so
        they've had that time to construct more expansive infrastructure
        compared.

------------------------------------------------------------------------

#### Step 6: Aggregate to County Level

Use spatial joins and aggregation to calculate county-level statistics
about vulnerable populations and hospital access.

**Your Task:**

```{r}
# Spatial join tracts to counties
vulnerable_tracts <- vulnerable_tracts %>%
  mutate(NAMELSADCO = str_remove_all(NAMELSADCO, " County"))

vulnerable_tracts$NAMELSADCO <- toupper(vulnerable_tracts$NAMELSADCO)

vulnerable_tracts <- st_transform(vulnerable_tracts, crs = 4326)

pa_counties <- st_transform(pa_counties, crs = 4326)

county_tracts <- st_join(vulnerable_tracts, pa_counties, left = TRUE)

# Aggregate statistics by county
aggregate_summary <- county_tracts %>%
  st_drop_geometry() %>%
  group_by(NAMELSADCO) %>%
  summarize(
    number_vulnerable = sum(vulnerable),
    number_underserved = sum(underserved),
    percent_underserved = (number_underserved / number_vulnerable) * 100,
    avg_nearby_hospital_dist = mean(nearest_hospital_mi),
    total_vulnerable_pop = sum(total_popE)
  ) %>%
  rename(
    county = NAMELSADCO
  )

pa_counties <- rename(pa_counties, county = COUNTY_NAM)

pa_mapping <- left_join(pa_counties, aggregate_summary, by = "county")

ggplot(pa_mapping) +
  geom_sf(aes(fill = percent_underserved), color = "white", size = 0.5) +
  scale_fill_viridis_c(
    name = "Percent Underserved",
    labels = comma,
    option = "magma"
  ) +
  labs(
    title = "Most Underserved Counties",
    subtitle = "Pennsylvania",
    caption = "Source: ACS 2019–2023"
  ) +
  theme_void()
```

**Required county-level statistics:**

-   Number of vulnerable tracts

-   Number of underserved tracts

-   Percentage of vulnerable tracts that are underserved

-   Average distance to nearest hospital for vulnerable tracts

-   Total vulnerable population

**Questions to answer:**

-   Which 5 counties have the highest percentage of underserved
    vulnerable tracts?

    -   Cameron (100%), Forest (100%), Sullivan (100%), Union (66.67%),
        and Clearfield (61.54%) counties.

-   Which counties have the most vulnerable people living far from
    hospitals?

    -   Counties with an average hospital distance of more than 15 miles
        and with the most vulnerable people in that group are Sullivan,
        Cameron, Forest, and Pike.

-   Are there any patterns in where underserved counties are located?

    -   It looks like more of the underserved counties are in the
        northern portion of Pennsylvania, and seemingly clustered toward
        the middle in addition. I also know those areas are peppered
        with several forested regions like national and state forests.

------------------------------------------------------------------------

#### Step 7: Create Summary Table

Create a professional table showing the top 10 priority counties for
healthcare investment.

**Your Task:**

```{r}
# Create and format priority counties table
kable_table <- pa_mapping[, c("county", "number_vulnerable",
                                "number_underserved", "percent_underserved",
                                "avg_nearby_hospital_dist", "total_vulnerable_pop")]

kable_table <- kable_table %>%
  mutate(
    avg_nearby_hospital_dist = drop_units(avg_nearby_hospital_dist)
  )

kable_table <- kable_table %>%
  st_drop_geometry() %>%
  rename(
    "County" = county,
    "Number of Vulnerable Tracts" = number_vulnerable,
    "Number of Underserved Tracts" = number_underserved,
    "Percent Underserved" = percent_underserved,
    "Average Mile(s) to Nearest Hospital" = avg_nearby_hospital_dist,
    "Total Vulnerable People" = total_vulnerable_pop
  )

kable_table <- kable_table %>%
  arrange(desc(`Percent Underserved`)) %>%
  slice(1:10)

kable(
  kable_table,
  digit = 2,
  format.args = list(big.mark = ","),
  caption = "<b>TOP 10 UNDERSERVED PENNSYLVANIA COUNTIES: For Priority Healthcare Investment</b>"
) %>%
  kable_styling(bootstrap_options = "striped") %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, color = "mistyrose", background = "red3")
```

**Requirements:**

-   Use `knitr::kable()` or similar for formatting

-   Include descriptive column names

-   Format numbers appropriately (commas for population, percentages,
    etc.)

-   Add an informative caption

-   Sort by priority (you decide the metric)

------------------------------------------------------------------------

## Part 2: Comprehensive Visualization

Using the skills from Week 3 (Data Visualization), create
publication-quality maps and charts.

### Map 1: County-Level Choropleth

Create a choropleth map showing healthcare access challenges at the
county level.

**Your Task:**

```{r}
# Create county-level access map
pa_hospitals <- st_transform(pa_hospitals, crs = 4326)
pa_mapping <- st_transform(pa_mapping, crs = 4326)

ggplot(pa_mapping) +
  geom_sf(
    aes(fill = percent_underserved),
    color = "white",
    size = 0.5
  ) +
  scale_fill_viridis_c(
    name = "Percent Underserved (%)",
    labels = scales::comma,
    option = "viridis"
  ) +
  geom_point(
    data = pa_hospitals,
    aes(x = LONGITUDE, y = LATITUDE, color = "Hospitals"),
    alpha = 0.75, inherit.aes = FALSE
  ) +
  scale_color_manual(
    name = "Hospitals",
    values = c("Hospitals" = "red")
  ) +
  labs(
    title = "MOST UNDERSERVED COUNTIES & HOSPITAL LOCATIONS",
    subtitle = "Pennsylvania, United States",
    caption = "Source: 5-Year American Community Survey (ACS) 2019–2023",
  ) +
  theme_void()
```

**Requirements:**

-   Fill counties by percentage of vulnerable tracts that are
    underserved

-   Include hospital locations as points

-   Use an appropriate color scheme

-   Include clear title, subtitle, and caption

-   Use `theme_void()` or similar clean theme

-   Add a legend with formatted labels

------------------------------------------------------------------------

### Map 2: Detailed Vulnerability Map

Create a map highlighting underserved vulnerable tracts.

**Your Task:**

```{r}
# Create detailed tract-level map
vulnerable_tracts <- st_transform(vulnerable_tracts, crs = 4326)
pa_tracts <- st_transform(pa_tracts, crs = 4326)

vulnerable_tracts <- vulnerable_tracts %>%
  rename("Underserved Tracts" = underserved)

choropleth_map <- ggplot(vulnerable_tracts) +
  geom_sf(
    aes(fill = `Underserved Tracts`),
  ) +
  geom_sf(
    data = pa_tracts, fill = NA,
    color = "midnightblue", linewidth = 0.35
  ) +
  geom_point(
    data = pa_hospitals,
    aes(x = LONGITUDE, y = LATITUDE, color = "Hospitals"),
    alpha = 0.75, inherit.aes = FALSE
  ) +
  scale_fill_manual(
    name = "Underserved Tracts",
    values = c("TRUE" = "gold2", "FALSE" = "lemonchiffon2"),
    labels = c("TRUE" = "Underserved", "FALSE" = "Vulnerable")
  ) +
  scale_color_manual(
    name = "Hospitals",
    values = c("Hospitals" = "red2")
  ) +
  labs(
    title = "UNDERSERVED AND VULNERABLE CENSUS TRACTS",
    subtitle = "With Hospital Locations in Pennsylvania, United States",
    caption = "Source: 5-Year American Community Survey (ACS) 2019–2023",
  ) +
  theme_void()

choropleth_map + theme(
  panel.background = element_rect(fill = "azure3", size = 0.75),
  legend.box.margin = margin(l = 20, unit = "pt"),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(family = "mono"),
  plot.title = element_text(face = "bold"),
  plot.subtitle = element_text(margin = margin(b = 20, t = 5, unit = "pt")),
  plot.caption = element_text(margin = margin(t = 10, unit = "pt"), face = "italic", family = "mono")
)
```

**Requirements:**

-   Show underserved vulnerable tracts in a contrasting color

-   Include county boundaries for context

-   Show hospital locations

-   Use appropriate visual hierarchy (what should stand out?)

-   Include informative title and subtitle

------------------------------------------------------------------------

### Chart: Distribution Analysis

Create a visualization showing the distribution of distances to
hospitals for vulnerable populations.

**Your Task:**

```{r}
# Create distribution visualization
vulnerable_tracts$nearest_hospital_mi <- drop_units(vulnerable_tracts$nearest_hospital_mi)

histogram_hospital <- ggplot(vulnerable_tracts, aes(x = nearest_hospital_mi)) +
  geom_histogram()

histogram_hospital_log <- ggplot(vulnerable_tracts, aes(x = log(nearest_hospital_mi))) +
  geom_histogram(
    color = "black",
    fill = "mistyrose2"
  ) +
  labs(
    title = "HISTOGRAM: NATURAL LOG OF DISTANCE TO NEAREST HOSPITAL",
    x = "Natural Log of Distance to Nearest Hospital",
    y = "Count") +
  theme_minimal() +
  theme(aspect.ratio = 1/3)

histogram_elderly <- ggplot(vulnerable_tracts, aes(x = age_65_overE)) +
  geom_histogram(
    color = "black",
    fill = "mistyrose2"
  ) +
  labs(
    title = "HISTOGRAM: VULNERABLE POPULATION OF AGES 65+",
    x = "Population of Ages 65+",
    y = "Count") +
  theme_minimal() +
  theme(aspect.ratio = 1/3)

scatter_plot <- ggplot(vulnerable_tracts, aes(x = log(nearest_hospital_mi), y = age_65_overE)) + 
  geom_point(
    aes(alpha = 0.5)) +
  geom_smooth(
    method = lm, linetype = "dashed",
    color = "red2", fill = "palegreen3") +
  labs(
    title = "SCATTER PLOT: HOSPITAL DISTANCE AND VULNERABLE POPULATION",
    subtitle = "Natural Log of Distance to Nearest Hospital vs. Population of Ages 65+\nwith Regression Line",
    caption = "Source: 5-Year American Community Survey (ACS) 2019–2023",
    x = "Natural Log of Distance to Nearest Hospital",
    y = "Population of Ages 65+") +
  theme(aspect.ratio = 4/3) +
  guides(alpha = "none")

faceted_plot <-
  (histogram_hospital_log + theme(
    plot.title = element_text(margin = margin(b = 10, unit = "pt"), face = "bold"),
    axis.title.x = element_text(margin = margin(t = 10, unit = "pt")),
    axis.title.y = element_text(margin = margin(r = 10, unit = "pt"))
  )) /
  (histogram_elderly + theme(
    plot.title = element_text(margin = margin(t = 20, b = 10, unit = "pt"), face = "bold"),
    axis.title.x = element_text(margin = margin(t = 10, unit = "pt")),
    axis.title.y = element_text(margin = margin(r = 10, unit = "pt"))
  )) /
  (scatter_plot + theme(
  legend.box.margin = margin(l = 20, unit = "pt"),
  legend.title = element_text(face = "bold"),
  legend.text = element_text(family = "mono"),
  plot.title = element_text(margin = margin(t = 20, unit = "pt"), face = "bold"),
  plot.subtitle = element_text(margin = margin(b = 10, t = 5, unit = "pt")),
  plot.caption = element_text(margin = margin(t = 10, unit = "pt"), face = "italic", family = "mono"),
  axis.title.x = element_text(margin = margin(t = 10, unit = "pt")),
  axis.title.y = element_text(margin = margin(r = 10, unit = "pt"))
  )) +
  plot_layout(heights = c(1, 1, 2))
```

```{r, fig.width = 8, fig.height = 12, dpi = 300}
faceted_plot
```

**Suggested chart types:**

-   Histogram or density plot of distances

-   Box plot comparing distances across regions

-   Bar chart of underserved tracts by county

-   Scatter plot of distance vs. vulnerable population size

**Requirements:**

-   Clear axes labels with units

-   Appropriate title

-   Professional formatting

-   Brief interpretation (1-2 sentences as a caption or in text)

    -   There's an upward regression line that seems to communicate and
        positive correlation between distance to the nearest hospital
        and the vulnerable population. However, a more in-depth
        statistical analysis will be needed to determine any
        significance, especially since the vast majority of points look
        like they are varied from the regression line. The natural log
        of the distance histogram was calculated due to the original
        data being highly right-skewed.

------------------------------------------------------------------------

## Part 3: Bring Your Own Data Analysis

Choose your own additional spatial dataset and conduct a supplementary
analysis.

### Challenge Options

Choose ONE of the following challenge exercises, or propose your own
research question using OpenDataPhilly data
(https://opendataphilly.org/datasets/).

**Note these are just loose suggestions to spark ideas - follow or make
your own as the data permits and as your ideas evolve. This analysis
should include bringing in your own dataset, ensuring the projection/CRS
of your layers align and are appropriate for the analysis (not lat/long
or geodetic coordinate systems). The analysis portion should include
some combination of spatial and attribute operations to answer a
relatively straightforward question**

------------------------------------------------------------------------

#### Education & Youth Services

**Option A: Educational Desert Analysis**

-   **Data:** Schools, Libraries, Recreation Centers, Census tracts
    (child population)

-   **Question:** "Which neighborhoods lack adequate educational
    infrastructure for children?"

-   **Operations:** Buffer schools/libraries (0.5 mile walking
    distance), identify coverage gaps, overlay with child population
    density

-   **Policy relevance:** School district planning, library placement,
    after-school program siting

**Option B: School Safety Zones**

-   **Data:** Schools, Crime Incidents, Bike Network

-   **Question:** "Are school zones safe for walking/biking, or are they
    crime hotspots?"

-   **Operations:** Buffer schools (1000ft safety zone), spatial join
    with crime incidents, assess bike infrastructure coverage

-   **Policy relevance:** Safe Routes to School programs, crossing guard
    placement

------------------------------------------------------------------------

#### Environmental Justice

**Option C: Green Space Equity**

-   **Data:** Parks, Street Trees, Census tracts (race/income
    demographics)

-   **Question:** "Do low-income and minority neighborhoods have
    equitable access to green space?"

-   **Operations:** Buffer parks (10-minute walk = 0.5 mile), calculate
    tree canopy or park acreage per capita, compare by demographics

-   **Policy relevance:** Climate resilience, environmental justice,
    urban forestry investment

------------------------------------------------------------------------

#### Public Safety & Justice

**Option D: Crime & Community Resources**

-   **Data:** Crime Incidents, Recreation Centers, Libraries, Street
    Lights

-   **Question:** "Are high-crime areas underserved by community
    resources?"

-   **Operations:** Aggregate crime counts to census tracts or
    neighborhoods, count community resources per area, spatial
    correlation analysis

-   **Policy relevance:** Community investment, violence prevention
    strategies

------------------------------------------------------------------------

#### Infrastructure & Services

**Option E: Polling Place Accessibility**

-   **Data:** Polling Places, SEPTA stops, Census tracts (elderly
    population, disability rates)

-   **Question:** "Are polling places accessible for elderly and
    disabled voters?"

-   **Operations:** Buffer polling places and transit stops, identify
    vulnerable populations, find areas lacking access

-   **Policy relevance:** Voting rights, election infrastructure, ADA
    compliance

------------------------------------------------------------------------

#### Health & Wellness

**Option F: Recreation & Population Health**

-   **Data:** Recreation Centers, Playgrounds, Parks, Census tracts
    (demographics)

-   **Question:** "Is lack of recreation access associated with
    vulnerable populations?"

-   **Operations:** Calculate recreation facilities per capita by
    neighborhood, buffer facilities for walking access, overlay with
    demographic indicators

-   **Policy relevance:** Public health investment, recreation
    programming, obesity prevention

------------------------------------------------------------------------

#### Emergency Services

**Option G: EMS Response Coverage**

-   **Data:** Fire Stations, EMS stations, Population density, High-rise
    buildings

-   **Question:** "Are population-dense areas adequately covered by
    emergency services?"

-   **Operations:** Create service area buffers (5-minute drive = \~2
    miles), assess population coverage, identify gaps in high-density
    areas

-   **Policy relevance:** Emergency preparedness, station siting
    decisions

------------------------------------------------------------------------

#### Arts & Culture

**Option H: Cultural Asset Distribution**

-   **Data:** Public Art, Museums, Historic sites/markers, Neighborhoods

-   **Question:** "Do all neighborhoods have equitable access to
    cultural amenities?"

-   **Operations:** Count cultural assets per neighborhood, normalize by
    population, compare distribution across demographic groups

-   **Policy relevance:** Cultural equity, tourism, quality of life,
    neighborhood identity

------------------------------------------------------------------------

### Data Sources

**OpenDataPhilly:** https://opendataphilly.org/datasets/

-   Most datasets available as GeoJSON, Shapefile, or CSV with
    coordinates

-   Always check the Metadata for a data dictionary of the fields.

**Additional Sources:**

-   **Pennsylvania Open Data:** https://data.pa.gov/

-   **Census Bureau (via tidycensus):** Demographics, economic
    indicators, commute patterns

-   **TIGER/Line (via tigris):** Geographic boundaries

### Recommended Starting Points

**If you're feeling confident:** Choose an advanced challenge with
multiple data layers. **If you are a beginner, choose something more
manageable that helps you understand the basics**

**If you have a different idea:** Propose your own question! Just make
sure:

-   You can access the spatial data

-   You can perform at least 2 spatial operations

### Your Analysis

**Your Task:**

1.  **Find and load additional data**
    -   Document your data source
    -   Check and standardize the CRS
    -   Provide basic summary statistics

```{r}
# Load your additional dataset
```

**Questions to answer:**

-   What dataset did you choose and why?

-   

-   What is the data source and date?

-   

-   How many features does it contain?

-   

-   What CRS is it in? Did you need to transform it?

-   

------------------------------------------------------------------------

2.  **Pose a research question**

Write a clear research statement that your analysis will answer.

**Examples:**

-   "Do vulnerable tracts have adequate public transit access to
    hospitals?"

-   "Are EMS stations appropriately located near vulnerable
    populations?"

-   "Do areas with low vehicle access have worse hospital access?"

------------------------------------------------------------------------

3.  **Conduct spatial analysis**

Use at least TWO spatial operations to answer your research question.

**Required operations (choose 2+):**

-   Buffers

-   Spatial joins

-   Spatial filtering with predicates

-   Distance calculations

-   Intersections or unions

-   Point-in-polygon aggregation

**Your Task:**

```{r}
# Your spatial analysis
```

**Analysis requirements:**

-   Clear code comments explaining each step

-   Appropriate CRS transformations

-   Summary statistics or counts

-   At least one map showing your findings

-   Brief interpretation of results (3-5 sentences)

**Your interpretation:**

\[Write your findings here\]

## Finally - A few comments about your incorporation of feedback!

Take a few moments to clean up your markdown document and then write a
line or two or three about how you may have incorporated feedback that
you received after your first assignment.

------------------------------------------------------------------------

## Submission Requirements

**What to submit:**

1.  **Rendered HTML document posted to your course portfolio** with all
    code, outputs, maps, and text
    -   Use `embed-resources: true` in YAML so it's a single file
    -   All code should run without errors
    -   All maps and charts should display correctly

**File naming:** `LastName_FirstName_Assignment2.html` and
`LastName_FirstName_Assignment2.qmd`

------------------------------------------------------------------------
